Faithful Variable Screening for\\ High-Dimensional Convex Regression

### 2015-11-19 tex/scam/reviews-v1/response-v1.tex

\item \rc{Intuitively, variable selection is very closely related to
estimating a function (or possibly its partial derivatives) in the
sup-norm.  It is well-known that accurate estimation of functions
(whether in $L^2$ or the sup-norm) with bounded smoothness in dimension
$s$ requires a sample size that grows exponentially in $s$.  For this
reason, the past result of Comminges and Dalalyan (suggesting an
exponential growth for variable selection) is to be expected.
However, in terms of metric entropy, the class of convex functions
behaves roughly like the class of two-smooth functions (e.g., Dryanov
(2009)).  So at first glance, this makes the result of this paper
rather surprising---namely, that one can somehow side-step any
exponential growth in $s$, even though the metric entropy is still
scaling as badly as the case of 2-smooth functions.  I wondered if the
authors could comment and shed insight on this fact, which seems a
little counterintuitive.  What I suspect that it could mean is that
the additive faithfulness (AF) condition, or one of the sufficient
conditions for AF (namely boundary flatness) are somehow very
restrictive.  Could the authors comment on these issues?}

Boundary flatness doesn't artificially make variable selection
easy. Boundary flatness is much more general than the product density,
and Comminges and Dalalyan showed that even under a product density
variable selection for general two-smooth functions is hard. Any
bounded density with compact support can be approximated arbitrary
well by a boundary flat density (see the newly added Example 3.2).

The intuition behind our result is the following. Start with the fact
that convex functions are additively faithful under a product
density. Next observe that the behavior of a convex function is constrained
by its behavior at the boundary; in particular, it is identically zero
if it's zero on the boundary. Therefore, it is sufficient if the 
underlying density resembles a product density at the boundary.  This is
exactly the notion that boundary flatness formalizes.

@@

\subsection*{Reviewer 1}

\rc{Conceptually, I have a few questions for the authors to address:}

\begin{enumerate}[(1)]
\item \rc{If the true regression function $f_0$ can actually be written as a
sum of univariate convex functions (plus noise), are there natural
conditions that one can provide for when additive faithfulness will
hold? Similarly, is there a simpler version of Theorem 3 that can be
stated when $f_0$ is actually a sum of univariate convex functions? This
would seem to depend on an appropriate statement for Assumption (A3).}

If the true regression function is a convex additive function, then
additive faithfulness holds---the best additive approximation
to an additive function is the function itself. The estimation
procedure becomes simpler.  In this case the DC stage is no longer necessary and can
be safely removed.

@@

To minimize predictive error, we can refit a non-additive convex
function after selecting a sparse model. Comparing the predictive
accuracy of the $f^*_k$s and $g^*_k$s against the non-additive true
regression function $f_0$ is beyond the scope of our paper, but we do
compare the finite sample estimate $\hat{f}_k$ and $\hat{g}_k$ against
the population level $f^*_k$ and $g^*_k$. Our analysis to
control false negatives also shows that $$\left| \mathbb{E}\left(
f_0(X) - \sum_{k=1}^p f^*_k(X_k) \right)^2 - \mathbb{E} \left( f_0(X)
- \sum_{k=1}^p \hat{f}_k(X_k))\right)^2 \right| \rightarrow 0$$ and
likewise for the $\hat{g}_k$s. Please see Theorem 8.3 and 8.4 in the
supplement.

We have added a paragraph in Section 3.4 that discusses prediction.

\item \rc{In order to improve clarity, the authors should define the convex
additive model at the top of page 5: $f(x_i) = \sum_{k=1}^p f_k(x_{ik})$, where the
$f_k$s are convex.  Otherwise, the notation in equation (2.3) is
initially a bit confusing.}

We have removed the old equation (2.3), and moved the discussion of convex
additive models to after AC/DC has been introduced, on page 6.

### 2015-11-30 tex/scam/experiments.tex
new support size plots

\subsubsection{Third simulation}
In the \textrm{third simulation} (Figure~\ref{Support}d,e), we use the
non-Gaussian distribution described above and set the covariance
to be $\Sigma_{ij}=\nu^{|i-j|}$ for $\nu = 0, 0.2, 0.5, 0.9$. We use the
non-additive $\Q$, same as in the first experiment, with $\alpha=0.5$
and fix $p=128, s=5$. We measure success not through exact recovery
but through faithful recovery. We say that a trial is successful if
(1) all relevant variables were recovered and (2) fewer than $20$
variables were marked as relevant overall (true sparsity $s=5$). We
use the same $\lambda$ as before. The probabilities of success are
computed from 40 independent trials and plotted against various values
of $\nu$ in Figure~\ref{Support}(d). Additionally, for $\nu = 0.5$, we
show the number of selected variables versus the sample size as a
box plot in Figure~\ref{Support}(e). As seen, for small to
moderate correlations, AC/DC can successfully recover the relevant
variables with only a small number of false positives.

In the fourth and fifth simulation, we use a softmax function as the ground truth,
\begin{align}
f_0(x_{iS}) = \log \left( \sum_{k=1}^K \exp( \beta_k^\tran x_{iS} ) \right) - \mu.
\end{align}
We generate random unit vectors as $\{\beta_k \in \R^s\}_{k=1,\ldots,K}$ and choose $\mu$ so that $f_0$ has mean-zero. We set $K = 7$ for all the experiments.

\subsubsection{Fourth simulation}
For the \textrm{fourth simulation} (Figure~\ref{Support}f,g), we let
$f_0$ be the softmax function and let $X$ be drawn from the boundary
flat mixture distribution described earlier with the Toeplitz
covariance $\Sigma_{ij}=\nu^{|i-j|}$ for $\nu = 0.5$. We set $s=5$ and
vary $p=128,256,512$. We use the same faithful recovery criteria as in
the third simulation and plot the probability of faithful recovery
against the number of samples in Figure~\ref{Support}(f). The
probabilities are computed over 30 independent trials. Also, for
$p=256$, we show the number of selected variables versus the sample
size as a box plot in Figure~\ref{Support}(g). The softmax function is
more challenging to estimate than the quadratic function; the softmax
function requires about $n > 1{,}500$ samples to achieve the same success
probability as the quadratic function with $n=1{,}000$. Regardless, we
see that increasing the ambient dimension $p$ does not significantly
affect the recovery probability.

@@

\begin{figure*}[!t]
\begin{center}
\begin{tabular}{cc}
%\hskip-10pt
\includegraphics[width=.42\textwidth]{figs/CurveT} &
%\hskip-10pt
\includegraphics[width=.42\textwidth]{figs/CurveR} \\
(a) quadratic $f_0$, independent $X$, varying $p$  & (b) quadratic $f_0$, independent $X$, varying $\Q$ \\
%\hskip-10pt
\includegraphics[width=.42\textwidth]{figs/Q} &
%\hskip-10pt
\includegraphics[width=.42\textwidth]{figs/CurveC}  \\
(c) four $\Q$ matrices used in (b) & (d) quadratic $f_0$, correlated $X$, varying correlation $\nu$ \\
\includegraphics[width=.42\textwidth]{figs/C_support_box} & 
\includegraphics[width=.42\textwidth]{figs/CurveS}\\
(e) recovered support size for $\nu=0.5$ in (d) & (f) softmax $f_0$, correlated $X$, varying $p$ \\
%\includegraphics[width=0.42\textwidth]{figs/S_support_box} & \\
%(g) recovered support size for $p=256$ in (f) & (h) softmax $f_0$, correlated $X$, varying $s$ 
\end{tabular}
\end{center}
\caption{Support recovery results.}
\label{Support}
\vskip10pt
\end{figure*}

>>>>

\subsubsection{Third simulation}
In the \textrm{third simulation} (figure~\ref{Support}d, figure~\ref{fig:support_size}f), we use the
non-Gaussian distribution described above and set the covariance
to be $\Sigma_{ij}=\nu^{|i-j|}$ for $\nu = 0, 0.2, 0.5, 0.9$. We use the
non-additive $\Q$, same as in the first experiment, with $\alpha=0.5$
and fix $p=128, s=5$. In figure~\ref{Support}d, we say that a trial is successful 
if all relevant variables were recovered, i.e., there are no false negatives. In figure~\ref{fig:support_size}f, we also show the total number of selected variables versus the sample size as boxplots; since the true sparsity level is 5, these plots show the number of false positives. We
use the same $\lambda$ as before. The probabilities of success are
computed from 40 independent trials and plotted against various values
of $\nu$ in Figure~\ref{Support}(d). As seen, for all correlation levels, AC/DC can successfully recover the relevant
variables with only a small number of false positives. For $\nu = 0, 0.2, 0.5$, the number of false positives steadily decrease with the sample size, but for $\nu = 0.9$, the number of false positives stays roughly constant. The latter phenomenon does not contradict our theory, which assumes $X_S \perp X_{S^c}$ for some set $S \supset S_0$, but it does demonstrate a weakness of our method when the design is highly correlated.

In the fourth and fifth simulation, we use a softmax function as the ground truth,
\begin{align}
f_0(x_{iS}) = \log \left( \sum_{k=1}^K \exp( \beta_k^\tran x_{iS} ) \right) - \mu.
\end{align}
We generate $K$ random unit vectors $\{\beta_k \in \R^s\}_{k=1,\ldots,K}$ and choose $\mu$ so that $f_0$ has mean-zero. We set $K = 7$ for all the experiments.

\subsubsection{Fourth simulation}
For the \textrm{fourth simulation} (figure~\ref{Support}e, figure~\ref{fig:support_size}g), we let
$f_0$ be the softmax function and let $X$ be drawn from the boundary
flat mixture distribution described earlier with the Toeplitz
covariance $\Sigma_{ij}=\nu^{|i-j|}$ for $\nu = 0.5$. We set $s=5$ and
vary $p=128,256,512$. We use the same faithful recovery criterion as in
the third simulation and plot the probability of faithful recovery
against the number of samples in Figure~\ref{Support}(e). The
probabilities are computed over 30 independent trials. Also, we show the number of selected variables versus the sample
size as boxplots in Figure~\ref{fig:support_size}(g). The softmax function is
more challenging to estimate than the quadratic function, regardless, we
see that increasing the ambient dimension $p$ does not significantly
affect the recovery probability. We also see from figure~\ref{fig:support_size}(g) that larger $p$ does not lead to more number of false positives.

@@

\begin{figure*}[!t]
\begin{center}
\begin{tabular}{cc}
%\hskip-10pt
\includegraphics[width=.42\textwidth]{figs/CurveT} &
%\hskip-10pt
\includegraphics[width=.42\textwidth]{figs/CurveR} \\
(a) quadratic $f_0$, independent $X$, varying $p$  & (b) quadratic $f_0$, independent $X$, varying $\Q$ \\
%\hskip-10pt
\includegraphics[width=.42\textwidth]{figs/Q} &
%\hskip-10pt
\includegraphics[width=.42\textwidth]{figs/CurveC}  \\
(c) four $\Q$ matrices used in (b) & (d) quadratic $f_0$, correlated $X$, varying correlation $\nu$ \\
%\multicolumn{2}{c}{\includegraphics[width=.8\textwidth]{figs/C_support_box}}\\
%\multicolumn{2}{c}{(e) recovered support size in (d)} \\
\multicolumn{2}{c}{\includegraphics[width=.42\textwidth]{figs/CurveS}} \\
\multicolumn{2}{c}{(e) softmax $f_0$, correlated $X$, varying $p$} \\
%\includegraphics[width=0.42\textwidth]{figs/S_support_box} & \\
%(g) recovered support size for $p=256$ in (f) & (h) softmax $f_0$, correlated $X$, varying $s$ 
\end{tabular}
\end{center}
\caption{Support recovery results.}
\label{Support}
\vskip10pt
\end{figure*}

\begin{figure*}
\begin{center}
\begin{tabular}{c}
\includegraphics[width=.84\textwidth]{figs/C_support_box}\\
(f) recovered support sizes for figure~\ref{Support}(d) \\
\includegraphics[width=0.84\textwidth]{figs/S_support_box}\\
(g) recovered support sizes for figure~\ref{Support}(e)
\end{tabular}
\end{center}
\caption{Recovered support size in correlated design cases}
\label{fig:support_size}
\end{figure*}
